### 2021-11-23 NSF2021/Description/specificaim1.tex
last changes to NSF submission

\subsubsection{Problem formulation}
%\begin{wrapfigure}[h!]
\begin{wrapfigure}[11]{r}{0.55\textwidth}
 \centering
 \vspace{-3pt}
 \includegraphics[width=0.55\textwidth]{figures/SpecificAim1/Domain.jpg}
 \caption{\label{fig:flow_map}
   In the HAP formulation, hydronamic interactions
   come from the Stokes equations and
   hydrophobic interactions come from the structure order parameter $u$.
   The colormap is for the intensity of $u$.}
\end{wrapfigure}

@@

The stress~\eqref{eq:stress} was first
introduced by~\cite{Fu2018_SIAM} and is the higher-dimensional
generalization of disjoining pressure \cite{MaRa76, ErLjCl89, KoNa15,
Nagle17, KUZMIN2005}. It is the mathematical ingredient responsible for
forming particle aggregates that sequester their hydrophobic surface
regions.

@@

\begin{wrapfigure}[14]{l}{2in}
  \vspace{-4pt}
  \centering
  \includegraphics[width=2in]{figures/SpecificAim1/OrderPDFs.jpg} 
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:OrderPDFs}
    Probability density functions for
    relative orientation of nearby particles
    for the data in 
    Figure~\ref{fig:self-assembly}, $t = t_{\text{final}}$.
  }
\end{wrapfigure}
Multilamelar bilayers arise when both sides of the particle
are hydrophobic (Figure~\ref{fig:self-assembly}(b)).
The boundary condition \eqref{eq:bcs}(b) 
gives a particle with a hydrophobic intensity that is greater
on the $\theta_i = 0$ side than on the $\theta_i = \pi$ side.
The initial self-assembly is similar to that in case (a).
The difference arises in the long-time dynamics where the bilayers
no longer remain well-separated. 
Rather, the bilayers form layers 
as a consequence of the interfacial tension of exposed particle
heads~\cite{Huetal19, deMeetal21}.

@@

\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.8in}m{1.8in}m{1.8in}}            
      (a)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{$t = t_{\text{final}}$}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:self-assembly}
    Rows (a)-(c) have the same initial configuration
    but have different boundary conditions $g.$ 
    In row (a), self-assembly with an amphiphilic boundary
    condition \eqref{eq:bcs}(a) results in a bilayer pattern shielding
    the hydrophobic core (yellow, $u > 0$) from the aqueous phase (blue, $u = 0$).
    Row (b) shows particles with a hydrophobic intensity that is
    greater on one side of the particle (magenta) than the other (yellow).
    In row (c), green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    Particles orient like sides to form a checkerboard pattern.\\
  }
\end{figure}

@@

Specific Aim 2 describes how we propose to deal with the nonlinear
equation~\eqref{eq:SL} numerically. Once we have established
implementations, we will consider further generalizations
of~\eqref{eq:free_energy}, for example to the Oseen-Frank energy for a
nematic liquid crystal~\cite{doi:10.1137/15M103844X,
Bartels,AIHPA_1997__66_4_411_0, deGennesProst,HaKiLi86,Virga}. For now,
Specific Aim 3 describes in detail how we plan to handle charges and
ions in solution.

>>>>

\subsubsection{Problem formulation}
%\begin{wrapfigure}[h!]
\begin{wrapfigure}[11]{r}{0.55\textwidth}
 \centering
 \vspace{-3pt}
 \includegraphics[width=0.55\textwidth]{figures/SpecificAim1/Domain.jpg}
 \caption{\label{fig:flow_map}
   In the HAP formulation,
   a system of exterior problems 
   defines the hydronamic and hydrophobic interactions.
   The colormap is for the intensity of $u$.
   The dynamics lead to self-organized structures like a vesicle bilayer.}
\end{wrapfigure}

@@

\begin{wrapfigure}[10]{r}{0.55\textwidth}
 \centering
 \vspace{-3pt}
 \includegraphics[width=0.55\textwidth]{figures/SpecificAim1/3Particles.jpg}
 \caption{\label{fig:3particles}
   \footnotesize
   Particles translate and rotate to lower the free energy $F[u]$
   subject to \eqref{eq:SL}, $f(u) = u^2$, and reach a minimal configuration in
   panel C. The colormap is blue for $u = 0$ and red for $u = 1$.
   The arrows are the velocity field $\mathbf{u}$.
 }
\end{wrapfigure}
The stress~\eqref{eq:stress} was first
introduced by~\cite{Fu2018_SIAM} and is the higher-dimensional
generalization of disjoining pressure \cite{MaRa76, ErLjCl89, KoNa15,
Nagle17, KUZMIN2005}. It is the mathematical ingredient responsible for
forming particle aggregates that sequester their hydrophobic surface
regions (Figure \ref{fig:3particles}).

@@

\begin{wrapfigure}[13]{l}{2in}
  \vspace{-4pt}
  \centering
  \includegraphics[width=2in]{figures/SpecificAim1/OrderPDFs.jpg} 
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:OrderPDFs}
    Probability density functions for
    relative orientation of nearby particles
    for the data in 
    Figure~\ref{fig:self-assembly}, long time.
  }
\end{wrapfigure}
Multilamelar bilayers arise when both sides of the particle
are hydrophobic (Figure~\ref{fig:self-assembly}(b)).
The boundary condition \eqref{eq:bcs}(b) 
gives a particle with a hydrophobic intensity that is greater
on the $\theta_i = 0$ side than on the $\theta_i = \pi$ side.
The initial self-assembly is similar to that in case (a).
The difference arises in the long-time dynamics where the bilayers
no longer remain well-separated. 
Rather, the bilayers form layers 
as a consequence of the interfacial tension of exposed particle
heads~\cite{Huetal19, deMeetal21}.

@@

\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.8in}m{1.8in}m{1.8in}}            
      (a)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{long time}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:self-assembly}
    Rows (a)-(c) have the same initial configuration
    but have different boundary conditions $g.$ 
    In row (a), self-assembly with an amphiphilic boundary
    condition \eqref{eq:bcs}(a) results in a bilayer pattern shielding
    the hydrophobic core (yellow, $u > 0$) from the aqueous phase (blue, $u = 0$).
    Row (b) shows particles with a hydrophobic intensity that is
    greater on one side of the particle (magenta) than the other (yellow).
    In row (c), green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    Particles orient like sides to form a checkerboard pattern.\\
  }
\end{figure}

@@

Specific Aim 2 describes how we propose to deal with the nonlinear
equation~\eqref{eq:SL} numerically. Once we have established
implementations, we will consider further generalizations
of~\eqref{eq:free_energy}, for example to the Oseen-Frank energy for a
nematic liquid crystal~\cite{doi:10.1137/15M103844X,
  Bartels,AIHPA_1997__66_4_411_0, deGennesProst,HaKiLi86,Virga}.
Another opportunity to build on the results of the proposed work,
is evaluating the role of particle multilayers on the stability of
pickering emulsions. Particle-particle interactions in the aqueous
layer (adjacent to the water-oil interface) may provide a
``reservoir'' of particles to stabilize the interface upon droplet deformation (see \cite{Bradley2016} and references therein).
For now,
Specific Aim 3 describes in detail how we plan to handle charges and
ions in solution.

### 2021-11-23 NSF2021/Description/summary.tex
last changes to NSF submission

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\pagestyle{empty}

\begin{document}
\begin{center}
{\bf \Large Summary}\\ Collaborative Research
\begin{tabular}{ccc}\\ 
Fordham University & Florida State University& New Jersey Institute of Technology\\
Rolf Ryham &            Bryan Quaife &              Yuan-Nan Young
\end{tabular}

\end{center}

\subsection*{Overview}
This proposal aims to advance mathematical modeling and analysis of collective dynamics of amphiphilic self-assembly.  Amphiphilic particles such as lipid molecules possess both hydrophobic and hydrophilic surfaces
that self-assemble into meso/macroscopic structures such as micelles and bilayers of lipids.  Experimentalists model self-assembly
using Janus particles, which have been widely used in soft matter physics as a model amphiphilic colloids and can be designed for catalytic activity or stimuli-responsive smart materials.

Amphiphilic self-assembly comes about due to a long-range attraction called the hydrophobic force.  This force is a major source of nonspecific interactions in biology and soft matter. Despite its ubiquity, theoretical developments explaining the basic underlying principles of the hydrophobic effect have come relatively late. Motivated by its broad applicability, the PIs recently developed a novel and intuitive mathematical model, called the hydrophobic attraction potential model, based on the linear response of water to surface perturbations. This model was used to show, for the first time, self-assembly into vesicle-like structures using a partial differential equation-based particle interaction.

The PIs propose to apply their hydrophobic attraction potential model to the self-organization dynamics of Janus particles.  They aim to generalize the linear response model to more general hydrophobic interactions, which will be numerically implemented using integral equation formulations.  Finally, they will investigate and simulate the hydrophobic interaction in the presence of electric charge and ions. The unifying aspects of the project are establishing a platform for efficiently simulating the collective dynamics at large scales, and the focusing on rigorous, mathematical analysis of the underlying model equations and complicated geometries.  The project supports undergraduate and graduate student collaborators.

\subsection*{Intellectual Merit}
The purpose of this research is to reach interesting physical phenomena with less computational cost than molecular dynamics, and account for more general features that continuum theory misses. The main ingredient is defining a nonlocal interaction through the solution of an elliptic boundary value problem that has the phenomenological characteristics of long-range hydrophobic attraction. This minimal model, while intuitive, is quite a general description of amphiphiles in solvent and gives rise to rich phenomena from Janus particle aggregates to correctly predicting elastic properties of bilayer. The technical research tasks include quantifying collective properties of amphiphilic ensembles, improving on mathematical models, efficient, high-order numerical algorithms for large-scale simulations, and incorporating external fields through electric charge.

\subsection*{Broader Impacts}
This project aims to advance the mathematical modeling of collective dynamics of amphiphilic particles. The simulations use a new, yet intuitive, approach that accounts for important and complex systems that are presently out of reach in computational material science. These complex systems include optimal shape design in metamaterials and fusion and fission of amphiphilic bilayer membranes. The models describing colloidal systems could be transformative in biomedicine and material science. The research draws from expertise in scientific computing, physics of fluids, and mathematics. The mathematical component incorporates leading variational techniques and offers insight into fundamentals of material science. The project brings socially consequential research into the classroom and offers 
undergraduates the opportunity to train alongside faculty and graduate students. With its combination of mathematical modeling, analysis, and scientific computing, the project highlights the importance of mathematics and computation to all areas of science and engineering.
\end{document}

>>>>

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\pagestyle{empty}
\linespread{1.05}
\begin{document}
\section*{Summary}

\subsection*{Overview}
\vspace{-0.1in}
This proposal aims to advance mathematical modeling and analysis of collective dynamics of amphiphilic self-assembly.  Amphiphilic particles such as lipid molecules possess both hydrophobic and hydrophilic surfaces and self-assemble into meso/macroscopic structures such as micelles and bilayers of lipids.  Experimentalists model self-assembly using Janus particles—spherical particles possessing a biphasic material label on either hemisphere.  Janus particles have been widely used in soft matter physics as a model amphiphilic colloid and can be designed for catalytic activity or stimuli-responsive smart materials.

Hydrophobic surfaces experience a nonadditive long-range attraction called the hydrophobic force.  This force is a major source of nonspecific interactions in biology and soft matter.  Despite its ubiquity, theoretical developments explaining its basic underlying principles have come relatively late.  Motivated by its broad applicability, the PIs recently developed a novel and intuitive mathematical model, called the hydrophobic attraction potential model, based on the linear response of water to surface perturbations. This model was used to show, for the first time, self-assembly of Janus particles into vesicle-like structures using a partial differential equation-based particle interaction.

The PIs propose to apply their hydrophobic attraction potential model to the self-organization dynamics of Janus particles.  They aim to generalize the linear-response model to more general hydrophobic interactions, and numerically implement these interactions using integral equation formulations.  Finally, they will investigate and simulate hydrophobic interactions in the presence of electric charge and ions. The unifying aspects of the project are establishing a platform for efficiently simulating the collective dynamics at large scales, and the focusing on rigorous, mathematical analysis of the underlying model equations and complicated geometries.  The project supports undergraduate and graduate student collaborators.

\subsection*{Intellectual Merit}
\vspace{-0.1in}
The purpose of this research is to reach interesting physical phenomena with less computational cost than molecular dynamics, and account for more general features that continuum theory misses. The main ingredient is defining a nonlocal interaction through the solution of an elliptic boundary value problem that has the phenomenological characteristics of long-range hydrophobic attraction. This minimal model, while intuitive, is quite a general description of amphiphiles in solvent and gives rise to rich phenomena from Janus particle aggregates to correctly predicting elastic properties of bilayer. The technical research tasks include quantifying collective properties of amphiphilic ensembles, improving on mathematical models, efficient, high-order numerical algorithms for large-scale simulations, and incorporating external fields through electric charge.

\subsection*{Broader Impacts}
\vspace{-0.1in}
This project aims to advance the mathematical modeling of collective dynamics of amphiphilic particles. The simulations use a new, yet intuitive, approach that accounts for important and complex systems that are presently out of reach in computational material science. These complex systems include optimal shape design in metamaterials and fusion and fission of amphiphilic bilayer membranes. The models describing colloidal systems could be transformative in biomedicine and material science. The research draws from expertise in scientific computing, physics of fluids, and mathematics. The mathematical component incorporates leading variational techniques and offers insight into fundamentals of material science. The project brings socially consequential research into the classroom and offers undergraduates the opportunity to train alongside faculty and graduate students. With its combination of mathematical modeling, analysis, and scientific computing, the project highlights the importance of mathematics and computation to all areas of science and engineering.
\end{document}
